On Optimality of Finite Element Integration

### 2015-12-03 TOMS_Paper/assembly.tex
minor fixes and todos

%If alternative (1) results better, the node \texttt{+} is marked as no longer expandable, which makes the recursion more efficient. Conversely,

%We then perform a second depth-first visit of the expression tree and we recursively label each node $n$ with a triple ${<}I_{jk},\ U_{jk},\ cost{>}$, where: $I_{jk}$ and $U_{jk}$ represent, respectively, the intersection and the union of the symbols depending on the multilinear loop nest (i.e., depending on test and trial functions) within $n$, whereas $cost$ denotes the operation count if generalized code motion were applied to all sub-expressions rooted in $n$ (i.e., if they were placed sufficiently far in the loop nest such that redundant computation were avoided). A crucial observation is that $I_{jk}$ and $U_{jk}$ are, in practice, always very small (a few units), even if the form is complex and includes several monomials. In addition, we note that $U_{jk}$ is tracked for the sole purpose of making the calculation of the parent $I_{jk}$ sets even faster.

%The algorithm starts with ``flattening'' the sub-expressions $\alpha_{eij}^{w} \beta_{eik}^{w}$ (by product expansion) such that all $L_j$- and $L_k$-dependent terms lie on the same level of the expression tree. We factorize, in sequence, $L_j$- ($L_k$-) and $L_k$- ($L_j$-) dependent terms. The choice of starting with $L_j$ or $L_k$ depends on two parameters: the number of unique terms, $n_{L}$, and the trip count, $tc_{L}$, of a loop $L$. It is fairly simple to see (and easily demonstrable by contradiction) that taking first $L$ such that $n_L \cdot tc_L$ is smallest leads to the largest operation count reduction. Assume, without loss of generality, that we start with $L_j$. Thanks to linearity, each factored product only has one symbol depending on $L_j$, and this symbol is now unique in the expression tree. The other sub-expression, independent of $L_j$, is, by definition, loop-invariant, and therefore hoisted such that redundant computation is avoided. Note that multilinearity ensures deterministic factorization, while perfectness ensures correctness of hoisting. We repeat the same process with $L_k$.
%
%For each of the expressions $\eta_{ei} = \sum_w \sigma_{ei}^{w}$ lifted at the level of $L_i$, we need understand whether applying a similar process -- that is, factorizing $L_i$-dependent symbols such that $L_e$-dependent sub-expressions can be hoisted -- actually results in reduced operation count (the problem described in Section~\ref{sec:se-rln}, point 3). Consider two sub-expressions $a_{ei}$ and $b_{ei}$ in $\eta_{ei}$, with $n$ and $m$ being the respective number of summands, and with $c$ being the number of summands in common. Given a product $a_{ei} b_{ei}$, the increase in operation count after product expansion is $n \cdot m - \binom c2$. Given a sum $a_{ei} + b_{ei}$, the operation count after factorization of common terms is $n + m - c$. We use these simple formulae, together with the trip count of $L_i$, to predict the impact -- and then decide on the application -- of sharing elimination at the level of $L_i$.
%
%In all this process, whenever a sub-expression is hoisted, we search and replace all common sub-expressions. This is probably useless for the actual operation count, since we expect any general-purpose compiler to be able to perform aggressive common sub-expression elimination\footnote{By inspection of assembly code, we have verified that this is actually the case with the Intel compiler v15.2.}, but it helps making the generated code much more readable.

% matrix notation, fixed normal form, so no need for fancy algors to do CSE
\end{Algo}

@@

Sharing elimination and pre-evaluation tackle the problem of minimizing spatial and temporal redundancy with a bounded working set in the loop nests and expressions defined in Section~\ref{sec:lnopt} \textit{using a very specific set of operators}. This must be kept in mind: obviously, the operation count might even be lower if we considered operators exploiting domain properties that we are here neglecting, such as redundancies in basis functions.

>>>>

%TODO sharing elimination of Li dependent stuff

%TODO CSE / CCSE /  SCSE In all this process, whenever a sub-expression is hoisted, we search and replace all common sub-expressions. This is probably useless for the actual operation count, since we expect any general-purpose compiler to be able to perform aggressive common sub-expression elimination\footnote{By inspection of assembly code, we have verified that this is actually the case with the Intel compiler v15.2.}, but it helps making the generated code much more readable.
\end{Algo}

@@

Sharing elimination and pre-evaluation tackle the problem of minimizing spatial and temporal redundancy given a set of memory constraints in finite element integration loop nests, using \textit{a very specific set of operators}. This needs be emphasized since, theoretically, one could find an even lower operation count by exploiting domain-specific properties, such as redundancies in basis functions.
